---
title: "MTCars Data Explorer"
author: "Marco Gigante"
date: "9/28/2025"
output: 
  ioslides_presentation:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(dplyr)
```

## Slide 1: MTCars Data Explorer
### Interactive Data Visualization Application

**Welcome to the MTCars Data Explorer!**

This Shiny application provides an interactive way to explore the classic MTCars dataset, allowing users to:

- Select different variables from the MTCars dataset
- Visualize data distributions through interactive histograms
- View comprehensive summary statistics
- Personalize their experience with custom greetings

**Built with:** R, Shiny, and ggplot2 for modern data visualization

## Slide 2: How It Works
### Simple Three-Step Process

**Step 1: Personalization**
- Enter your name in the text input field
- Receive a personalized greeting

**Step 2: Variable Selection**
- Choose from 11 different MTCars variables:
  - Miles per Gallon, Cylinders, Displacement
  - Horsepower, Weight, Transmission type, and more

**Step 3: Explore Results**
- View interactive histogram with professional styling
- Examine detailed summary statistics including mean, median, quartiles
- All outputs update automatically based on your selections

## Slide 3: Live Demo
### MTCars Data in Action

```{r demo-plot, echo=TRUE, fig.height=4}
# Plotting Miles per Gallon distribution from MTCars dataset
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(bins = 10, 
                 fill = "steelblue", 
                 color = "white", 
                 alpha = 0.7) +
  labs(title = "Distribution of MPG in MTCars Dataset",
       x = "Miles per Gallon",
       y = "Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14))
```
---
**Real-time statistics:** Mean = `r round(mean(mtcars$mpg), 2)`, Median = `r median(mtcars$mpg)`, Range = `r min(mtcars$mpg)` - `r max(mtcars$mpg)`
